Flexible Probabilistic Forecasting with Distributional Trees and Forests

### 2018-04-04 pkg/disttree/inst/MIP-Seminar/slides.Rnw
revised slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2125 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: fbe982d0fbe31324ad2cb9e32e15b077fefcc8b5

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are availabe (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

@@

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\visible<2,3, 6-> {Assumption:}
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2,3, 6-> {\node[] at (0,0.4) {X};}
\visible<4-5>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {15.03.2018};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2,3, 6->{
\node[] at (3.5,0.4) {Y};
}
\visible<4-5>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {16.03.2018};
}
\end{tikzpicture}
\end{center}
\visible<3, 6->{
\vspace{-2cm}
Approximation of reality based on a set of observations.\\
}
\visible<5>{
\vspace{2cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums\\
}

@@

\begin{frame}[fragile]
\frametitle{Data modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$.\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

@@

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
nobs <- 120
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- ytree_up <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.6) 4 else 9
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 4 else 9
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4 else 9
ytree_up <- rep.int(15, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.6) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
#%for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Algorithmic modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3-6>{Tree model\only<3>{\\}\only<4-6>{s}}\only<7->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-0.96cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain today, humidity\\
Y = rain tomorrow}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
}
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\vspace{0.1cm}

@@

mtext("today", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])

@@

mtext("today", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain tomorrow", side = 2, las = FALSE, line = 1, cex = 3)

@@

lines(x = x, y = ytree_up, col = pal["tree"], lwd=7)

@@

lines(x = x, y = ytree_up, col = "blue", lwd=7)

@@

lines(x = x, y = ytree_up, col = "orange", lwd=7)

@@

@@

@@

\begin{frame}
\frametitle{Goals}
\vspace{-0.7 cm}
\begin{minipage}{0.7\textwidth}
\textbf{Available:} \\
\vspace{-0.3 cm}
\begin{itemize}
  \item<2-> Probabilistic forecast\\
  $\rightarrow$ Distributional approach
  \item<3-> Automatic detection of steps and abrupt changes,
  automatic selection of variables and interactions\\
  $\rightarrow$ Tree approach
  \item<4-> Smoothing and regularization\\
  $\rightarrow$ Forest approach
\end{itemize}
\end{minipage}
<<motivation_GAMLSS_syn, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
nobs <- 200
x_g <- c(1:nobs)/nobs
x_g <- 2*(x_g-0.5)
ytrue_g <- x_g^3
var <- exp(-(2*x_g)^2)/2
y_g <- ytrue_g + rnorm(nobs, 0, 0.1 + var)
@
<<motivation_regtree_syn, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
nobs <- 120
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- ytree_up <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.6) 4 else 9
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 4 else 9
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4 else 9
ytree_up <- rep.int(15, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.6) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
\begin{minipage}{0.25\textwidth}
\vspace{0.7cm}
\visible<2->{
<<plot_motivation_GAMLSS_syn, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,0,3,0))
plot(x_g, y_g, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x_g, rev(x_g)), c(ytrue_g + 0.1 + var, rev(ytrue_g - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x_g, ytrue_g, col = pal["forest"], lwd=12)
points(x_g, y_g, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\visible<3->{
<<plot_motivation_regtree_syn, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,0,3,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=12)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=12, lty = 2)
@
}
\visible<4->{
<<plot_motivation_randforest4_syn, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,0,3,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=12, lty = 2)
lines(x = x, y = yforest, col = pal["forest"], lwd=12)
@
}
\end{minipage}

%\vspace{0.1cm}

\visible<5->{
\textbf{Missing:} Synthesis\\
}

\vspace{0.5cm}

\visible<6->{
$\Rightarrow$ PhD-Thesis  
}
\visible<7->{
(what I've been working on for the last 2 years)
}
\end{frame}

>>>>

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are available (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

@@

\begin{frame}[fragile]
\frametitle{Statistical Modeling}
\visible<2,3, 6-> {Assumption:}
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2,3, 6-> {\node[] at (0,0.4) {X};}
\visible<4-5>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {15.03.2018};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2,3, 6->{
\node[] at (3.5,0.4) {Y};
}
\visible<4-5>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {16.03.2018};
}
\end{tikzpicture}
\end{center}
\visible<3, 6->{
\vspace{-2cm}
Approximation of reality based on a set of observations.\\
}
\visible<5>{
\vspace{2cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums\\
}

@@

\begin{frame}[fragile]
\frametitle{Data Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$.\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

@@

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
nobs <- 120
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- 
  ytree_up <- ytree_up2 <- ytree_up3 <- ytree_up4 <- 
  yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.63) 4.3 else 9.4
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 3.7 else 9.2
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4.4 else 8.7
ytree_up <- rep.int(15, length(x))
ytree_up2 <- rep.int(14.5, length(x))
ytree_up3 <- rep.int(15.3, length(x))
ytree_up4 <- rep.int(15.5, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.63) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
#%for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Algorithmic Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3>{Tree model}\only<4-10>{Subsampling}\only<3>{\\}\only<11->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-0.96cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain now, humidity\\
Y = rain 24h}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
}
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\vspace{0.1cm}

@@

mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])

@@

mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)

@@

lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)

@@

lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up3, col = "blue", lwd=7)

@@

@@

lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up3, col = "blue", lwd=7)

@@

lines(x = x, y = ytree_up4, col = "orange", lwd=7)

@@

\begin{frame}
\frametitle{Goals}
\vspace{-0.4cm}
\begin{minipage}{0.7\textwidth}
\textbf{Available:} \\
\vspace{-0.5 cm}
\begin{itemize}
  \item Distributional regression:\\
  Probabilistic forecast.
  \item Regression Trees:\\
  Automatic detection of abrupt changes,
  and selection of variables/interactions.
  \item Random Forests:\\
  Smoothing and regularization.
\end{itemize}

\vspace{1cm}

\visible<2->{
\textbf{Missing:} Synthesis

\vspace{0.2cm}

$\Rightarrow$ PhD-Thesis

%{\small 
(what I've been working on the last 2 years)
}
\end{minipage}
<<motivation_GAMLSS_syn, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
nobs <- 200
x_g <- c(1:nobs)/nobs
x_g <- 2*(x_g-0.5)
ytrue_g <- x_g^3
var_g <- exp(-(2*x_g)^2)/2
y_g <- ytrue_g + rnorm(nobs, 0, 0.1 + var_g)
@
<<motivation_regtree_syn, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(74)
nobs <- 170
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- ytree_up <- yforest <- var <- var_up <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.2) 4 else 4+5*plogis(kappa*(3*(x[i]-0.33)-1))
ytrue_up <- rep.int(15, length(x))
for(i in 1:nobs) var_up[i] <- if(x[i]<0.15) 3 else (if(x[i]<0.35) 2 else (if(x[i]<0.5) 2 else (if(x[i]<0.7) 1 else (if(x[i]<0.8) 1.5 else (if(x[i]<0.9) 2 else (2.5))))))
for(i in 1:nobs) var[i] <- if(x[i]<0.15) 4 else (if(x[i]<0.2) 3 else (if(x[i]<0.25) 2 else (if(x[i]<0.4) 1 else (if(x[i]<0.8) 2 else (if(x[i]<0.9) 2.5 else (3))))))
y <- ytrue + rnorm(nobs,0, sqrt(var))
y_up <- ytrue_up + rnorm(nobs,0, sqrt(var_up))
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
ytree_up <- rep.int(15, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.6) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
\begin{minipage}{0.25\textwidth}
\vspace{0.4cm}
<<plot_motivation_GAMLSS_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x_g, y_g, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x_g, rev(x_g)), c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g)),
  col = pallight["forest"], border = "transparent")
lines(x_g, ytrue_g, col = pal["forest"], lwd=10)
points(x_g, y_g, col = "slategray", pch = 19)
box(lwd = 5)
@

<<plot_motivation_regtree_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=10)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
@

<<plot_motivation_randforest4_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@

\vspace{0.2cm}

\visible<2->{
<<plot_motivation_distforest, fig=TRUE, echo=FALSE, width=6>>=
par(mar=c(0,0,3,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
polygon(c(x, rev(x)), c(yforest + var, rev(yforest - var)),
  col = pallight["forest"], border = "transparent")
polygon(c(x, rev(x)), c(ytrue_up + var_up, rev(ytrue_up - var_up)),
  col = pallight["forest"], border = "transparent")
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@
}
\end{minipage}
\end{frame}
